\LARGE \bf Enhancing Security in Multi-Robot Systems through Co-Observation Planning on Unsecured Trajectories

### 2023-09-01 paper.tex
Updates from Overleaf

\input{preamble/common}
\newcommand{\rrtstar}{$\texttt{RRT}^\texttt{*}$}

@@

We formulate the planning problem into an optimal trajectory optimization problem to minimize arbitrary smooth objective functions. The security constraints including co-observation schedule and reachability analysis, alongside with traditional path planning constraints, are transformed into different types of spatio-temporal constraints in the optimization problem. We denote as $q_{ij}\in\real{m}$ the position of agent $i$ at the discrete-time index $j$, with $m$ representing the dimension of the state space. For a team of $n$ agents, and a task time horizon of $T$, the overall trajectory of the multi-agent system can be represented as an aggregated vector $\vq\in \real{n m T}$. 
The goal of our path planning problem is to minimize or maximize an objective function $\varPhi(\vq)$ under a set of nonlinear constraints described by a set $\Omega$, formally:
\begin{equation}\label{eq:general-problem}
	\begin{split}
		\min/\max & \quad \varPhi(\vq)\\
		\textrm{subject to} &\quad \vq \in \Omega.
	\end{split}
\end{equation}

As an example application, we assume that there exists an underlying slowly-time-varying scalar or vector field $\vx$ whose value needs to be estimated at a given number of discrete locations. To explore unknown spatially-distributed fields, autonomous agents are usually required to traverse the entire unknown environment while collecting sensory data to estimate a corresponding map (Figure~\ref{fig:ADMM-example-application}). We model the map as a set of locations of interests arranged on a regular grid, where at each location, we assume that there is a corresponding slowly-time-varying quantity. For our purposes, we are interested in finding paths that best reconstruct the field, i.e., that achieve the minimum uncertainty. We define the cost function $\varPhi(\vq)$ to be an approximation of the uncertainty at any of the given points, and we aim to minimize such maximum uncertainty. We focus on a particular choice of objective function $\varPhi(\vq)$, in the context of a mapping and estimation applications, for the optimization problem \eqref{eq:general-problem}. \zyang{Do we need detail of the problem formulations?}

@@

\subsection{Preliminary}
In this section, we review various mathematical concepts that will provide the foundations and context for the solver.

@@

We consider the following types of traditional path planning constraints in the ADMM constraints:
\subsubsection*{Velocity constraints}
  The movement of each agent is constrained by a maximum distance in any direction over a single discrete time step. We then define the function $D(\vq)$ to return the velocity vectors for the $i$-th agent at time step $j$:
\begin{equation}
     D_{ij}(\vq) = q_{ij}-q_{i(j-1)}, \quad j\in\{1,\ldots, T\}
\end{equation}
and the constraint set is:
\begin{equation}\label{eq:velocity_constraint}
   \sZ_{ij} =\bigl\{z \mid \norm{z} \leq v_{max}\bigr\}.
\end{equation}
The projection operator $\Pi_\zeta(z)$ for this constraint  is simply a projection of $z$ to the interior of a sphere centered at the origin and having radius $v_{max}$. This constraint could be adapted to enforce more refined dynamical models.

@@

 \subsubsection{Householder rotations}\label{sec:householder}

  To formulate the operator $\Pi_\sZ$ for our novel constraint, we will define a differentiable transformation of the constraint set to a canonical form. This transformation will include a rotation that we derive from a modified version of Householder transformations \cite{householder1958unitary}. With respect to the standard definition, our modification ensures that the final operator is a proper rotation (i.e., not a reflection). We call our version of the operator a \emph{Householder rotation}. In this section we derive Householder rotations and their differentials for the 3-D case; the 2-D case can be easily obtained by embedding it in the $z=0$ plane.

@@

\subsection{Definition of reachability constraint}\label{sec:reachability}
  In this section, we first provide the definition of an \emph{ellipsoidal reachability region}. We then provide a differentiable map to transform such region in a canonical axis-aligned form where the operator $\Pi_\zeta$ and its differential can be obtained; these operators are then extended to the general case via the aforementioned transform. The overall goal is to define the functions $D(q)$, its differential, and the operator $\Pi_\zeta$ for \emph{ellipsoidal reachability regions} with respect to the forbidden regions that can then be used in the ADMM formulation in section \ref{chapter:ADMM review}.

The reachability region is defined as the set of locations $q(t)$ that a robot can reach between two given fixed positions:

@@

 \begin{figure}
    \centering
    \includegraphics[width=0.6\linewidth]{Reachability}
    \caption{The ellipse showcases the reachability region. The black line is the planned trajectory of a robot, $q_1$ and $q_2$ are two locations this robot are expected at given time $t_1$ and $t_2$, blue and red line are possible trajectories if the robot is compromised in after reaching $q_1$. Robot that goes outside the reachability region to forbidden region can only reach $q_{2}'$ instead of $q_{2}$ at time $t_{2}$, even if it drives directly toward $q_{2}$ at full speed $v_{max}$ after the intrusion into forbidden region.}
    \label{fig:EllipseConstraintExample}
  \end{figure}

@@

\subsection{Reachability constraints using ellipsoidal constraints and their differential}\label{sec:ellipoid constraints}
In later sections, we define constraint formulation of the ellipsoid regions against different types of forbidden regions: a point, a plane, a segment, and a convex polygon. For each one, our goal is to define the function $D(q)$, the set $\zeta$ and the projection $\Pi_\zeta$ that can be incorporated in the ADMM optimization (\ref{eq:ADMMSetConstraint_modified}); we also include derivations for the differential $\partial_qD(q)$, which can be used to significantly speed up the optimization.

@@

\begin{equation}\label{eq:ellipse-point-projection}
\begin{split}
\argmin_{q_p} &\quad \norm{q_{avoid}-q_p}^2 \\
s.t. &\quad q_p\in \cE^{c}.
\end{split}
\end{equation}
where $\cE^{c}$ is the set complement of the region $\cE$.

@@

  \subsubsection{Projection to the standard ellipse}
  The ellipse $\cE$ expressed in $\cF_\cE$ is given by $\Eframe{\cE}=\{\Eframe{q}\in\real{m}:d(q^\cE_1,\Eframe{q})+d(\Eframe{q},q^\cE_2)<2a\}$, where the coordinates of the two foci $q^\cE_1,q^\cE_2$ in $\cF_\cE$ are
  \begin{align}
    q^\cE_1=\bmat{c & 0 & 0}\transpose, && q^\cE_2=\bmat{-c & 0 & 0}\transpose,
  \end{align}
  with $c=\frac{\norm{q_2-q_1}}{2}$.

@@

For a intermediate step to get the relationship between a polygon shaped forbidden region, we first consider a forbidden region in the shape of a hyperplane $\mathcal{L}(\tilde{q}) = \{\tilde{q}\in\mathbb{R}^n: \vn\transpose \tilde{q} = d\}$ (as shown in Figure \ref{fig:Ellipse-to-plane}). The reachability constraint now can then be defined as $\mathcal{L} \cap \mathcal{E}(q_1,q_2,a) = \emptyset$.
  Using the transformation introduced in (\ref{sec:rotation2Standard}), the ellipse can be transferred into a standard one with the new hyperplane in the form of $\Eframe \cL(\tilde{q}) = \{  \tilde{q}\in\mathbb{R}^m:  \vn_\cE\transpose \tilde{q} = d_{\cE}\}$, where,
  \begin{equation}
    n_\cE = H(q_1,q_2)n
\end{equation}
\begin{equation} 
	d_\cE =-\vn\transpose o + d
  \end{equation}
\subsubsection{Projection via the tangent interpolation point}
  For each $\cL_\cE$, there exist two planes $\Eframe \cL_{1} = \{\tilde{q}\in\mathbb{R}^m: \vn_\cE \transpose \tilde{q} =  d_{\cE t} \}$ and $\Eframe \cL_{2} = \{\tilde{q}\in\mathbb{R}^m: \vn_\cE \transpose \tilde{q} = - d_{\cE t} \}$ that are parallel to $\cL$ and tangent to the ellipse, where
  \begin{equation}
    d_{\cE t} = \sqrt{ n_\cE\transpose Q^{-1} n_\cE}.
  \end{equation}
  The corresponding tangent points are given as 
  \begin{equation}
    p^\cE_{1} = \frac{  d_{\cE t} Q^{-1}   n_\cE}{ n_\cE \transpose Q^{-1}  n_\cE} = \frac{Q^{-1} n_\cE}{ d_{\cE t}},
  \end{equation}
  \begin{equation}
   p^\cE_{2} = -  p^\cE_{1}.
  \end{equation}
  We use these two tangent points to define a novel measure of displacement between a plane and an ellipse.
  \begin{definition}
    The \emph{tangent interpolation point} is defined as 
    \begin{equation}\label{equ:ProjectPoint}
      p^\cE_{\cL} = \frac{ d_\cE Q^{-1} n_\cE}{ n_\cE\transpose Q^{-1} n_\cE}
    \end{equation}
  \end{definition}
  Note that $p^\cE_{\cL} \in \mathcal{L}$ and when  $d_\cE=d_{\cE t}$ or $d_\cE=-d_{\cE t}$, $p^\cE_{\cL}=p^\cE_{1}$ or $p^\cE_{\cL}=p^\cE_{2}$, respectively. 
  % it is easy to see that $\Eframe p_{\cL} \in \mathcal{L}$, and $p_{\mathcal{L}}$ can be used to indicate the relationship between the $\mathcal{L}$ and $\mathcal{E}$, which is shown below. \zyang{how to emphaise this? maybe a graph here?}
  %The reason such that a specific point is used is for the convenience of evaluating ellipse-plane-segment relationship in Section~\ref{chapter:ellipse to segment}.
  It is clear that when $ d_\cE \in [- d_{\cE t},  d_{\cE t}]$, the plane $\cL$ and the ellipsoid $\cE$ have at least one intersection, thus violating our desired reachability constraint. Using this fact, we define the following projection operator:
  \begin{equation}\label{equ:plane2ellipse}
    \pi^\cE_{\vn \cE}(q) =\begin{cases}
      p^\cE_{t1}-p^\cE_{\cL} &  d_\cE \in [0, d_{\cE t}], \\
      p^\cE_{t2}-p^\cE_{\cL} &  d_\cE \in [- d_{\cE t},0), \\
      {0} & \textrm{otherwise}.
    \end{cases}
  \end{equation}
  \subsubsection{ADMM constraints}
  Transforming the projection $\pi^\cE$ back to $\cF$ we have:
  \begin{equation}\label{equ:constraint plane to ellipse}
    D_{\vn}(q)= H^{-1}(q(t_1),q(t_2))\pi^\cE_{\vn \cE}(q)+o,
  \end{equation}
  with the corresponding constraint written as
  \begin{equation}
    \zeta_\vn = \{q\in\mathbb{R}^{nm}:\norm{D_{\vn}(q)} = 0 \}
  \end{equation}
  \begin{equation}
   \quad\Pi_\vn(D_{\vn}(q)) = {0}
  \end{equation}

@@

\begin{figure}[htbp]
\begin{center}
\includegraphics[width=0.6\linewidth]{Ellipse2safezone_V2}
\caption{Polygen-ellipse constraint. This constraint is simplified to either a plane-ellipse constraint or a point-ellipse constraint.}
\label{fig:Ellipse-to-safezone}
\end{center}
\end{figure}

>>>>

\input{preamble/common}
\newcommand{\rrtstar}{$\texttt{RRT}^\texttt{*}$}
\newtheorem{constraint}{ADMM constraint}
\newtheoreom{example}{Example}

@@

We formulate the planning problem into an optimal trajectory optimization problem to minimize arbitrary smooth objective functions. The security constraints, including those from the co-observation schedule, reachability analysis, and the traditional path planning constraints, are transformed into different types of spatio-temporal sets in the optimization problem. We denote as $q_{ij}\in\real{m}$ the position of agent $i$ at the discrete-time index $j$, with $m$ representing the dimension of the state space. For a team of $n$ agents, and a task time horizon of $T$, the overall trajectory of the multi-agent system can be represented as an aggregated vector $\vq\in \real{n m T}$. 
The goal of our path planning problem is to minimize or maximize an objective function $\varPhi(\vq)$ under a set of nonlinear constraints described by a set $\Omega$, formally:
\begin{equation}\label{eq:general-problem}
	\begin{split}
		\min/\max & \quad \varPhi(\vq)\\
		\textrm{subject to} &\quad \vq \in \Omega.
	\end{split}
\end{equation}

To give a concrete example of the cost $\varPhi$ and the set $\Omega$, we introduce a representative application that will be used for all the simulations throughout the paper.
\begin{example} Assume that there exists an underlying slowly-time-varying scalar or vector field $\vx$ whose value needs to be estimated at a given number of discrete locations. To explore unknown spatially-distributed fields, autonomous agents are usually required to traverse the entire unknown environment while collecting sensory data to estimate a corresponding map (Figure~\ref{fig:ADMM-example-application}). We model the map as a set of locations of interests arranged on a regular grid, where at each location, we assume that there is a corresponding slowly-time-varying quantity. For our purposes, we are interested in finding paths that best reconstruct the field, i.e., that achieve the minimum uncertainty. 
\rtron{Define $\varPhi$ and $\Omega$ in words but briefly}
%We define the cost function $\varPhi(\vq)$ to be an approximation of the uncertainty at any of the given points, and we aim to minimize such maximum uncertainty. We focus on a particular choice of objective function $\varPhi(\vq)$, in the context of a mapping and estimation applications, for the optimization problem \eqref{eq:general-problem}. \zyang{Do we need detail of the problem formulations?}
\end{example}

@@

\subsection{Preliminaries}
In this section, we review various mathematical concepts that will provide the foundations and context for the solver.

@@

We consider the following types of traditional path planning constraints in the ADMM constraints:
\subsubsection*{Velocity constraints}
  The movement of each agent is constrained by a maximum distance in any direction over a single discrete time step. We then define the function $D(\vq)$ to return the velocity vectors for the $i$-th agent at time step $j$; its result then needs to be constrained to a sphere.

  \begin{constraint}[Velocity constraint]
\begin{align}
     D_{ij}(\vq) &= q_{ij}-q_{i(j-1)}, \quad j\in\{1,\ldots, T\}\\
\label{eq:velocity_constraint}
   \sZ_{ij} &=\bigl\{z \mid \norm{z} \leq v_{max}\bigr\}\\
   \Pi_{\sZ_{ij}}(z)&=\begin{cases}
   v_{\max}\frac{z}{\norm{z}} \textrm{ if } \norm{z}>v_{\max}\\
   z\textrm{ otherwise}
   \end{cases}
\end{align}
\end{constraint}
\rtron{Put all the constraints using this $\LaTeX$ environment.}
\rtron{Make sure notation for $\sZ$ is consistent}
%The projection operator $\Pi_\zeta(z)$ for this constraint  is simply a projection of $z$ to the interior of a sphere centered at the origin and having radius $v_{max}$. This constraint could be adapted to enforce more refined dynamical models.

@@

 \subsubsection{Householder rotations}\label{sec:householder}
 \rtron{Move to right after the definition of differential}
  To formulate the operator $\Pi_\sZ$ for our novel constraint, we will define a differentiable transformation of the constraint set to a canonical form. This transformation will include a rotation that we derive from a modified version of Householder transformations \cite{householder1958unitary}. With respect to the standard definition, our modification ensures that the final operator is a proper rotation (i.e., not a reflection). We call our version of the operator a \emph{Householder rotation}. In this section we derive Householder rotations and their differentials for the 3-D case; the 2-D case can be easily obtained by embedding it in the $z=0$ plane.

@@

\subsection{Definition of ellipsoidal reachability regions}\label{sec:reachability}
  In this section, we define  \emph{ellipsoidal reachability regions} based on pairs of locations on a trajectory, and a transformation of such region in axis-aligned form. These definitions will be used to define different types of reachability constraints (between an ellipsoid and different geometric entities such as a point, line, line segment, and polygon), their projections, and differentials in subsequent sections. 
  %differentiable map to transform such region in a canonical axis-aligned form where the operator $\Pi_\zeta$ and its differential can be obtained; these operators are then extended to the general case via the aforementioned transform. The overall goal is to define the functions $D(q)$, its differential, and the operator $\Pi_\zeta$ for \emph{ellipsoidal reachability regions} with respect to the forbidden regions that can then be 
  These will be used in the ADMM formulation in section \ref{chapter:ADMM review}.

%The reachability region is defined as the set of locations $q(t)$ that a robot can reach between two given fixed positions:

@@

 \begin{figure}
    \centering
    \includegraphics[width=0.6\linewidth]{Reachability}
\rtron{Extend the figure to give a graphical representation of the canonical transformation}
    \caption{The ellipse showcases the reachability region. The black line is the planned trajectory of a robot, $q_1$ and $q_2$ are two locations this robot are expected at given time $t_1$ and $t_2$, blue and red line are possible trajectories if the robot is compromised in after reaching $q_1$. Robot that goes outside the reachability region to forbidden region can only reach $q_{2}'$ instead of $q_{2}$ at time $t_{2}$, even if it drives directly toward $q_{2}$ at full speed $v_{max}$ after the intrusion into forbidden region.}
    \label{fig:EllipseConstraintExample}
  \end{figure}

@@

%\subsection{Reachability constraints using ellipsoidal constraints and their differential}\label{sec:ellipoid constraints}
\rtron{Make sure that, throughout the paper, we use reachability ellipsoid instead of ellipsoidal regions}
We are now ready to formulate constraints based on reachability ellipsoids against different types of forbidden regions: a point, a plane, a segment, and a convex polygon. For each one, our goal is to define the function $D(q)$, its differential $\partial_qD(q)$, the set $\zeta$ and the projection $\Pi_\zeta$ that can be incorporated in the ADMM optimization (\ref{eq:ADMMSetConstraint_modified}).

@@

\rtron{For these constraints, add $\pi$ inside the constraint environment, then follow that environment with the proposition with the differential of $\pi$.}

\begin{equation}\label{eq:ellipse-point-projection}
\pi_{p\cE}=\argmin_{q_p\in\cE^c} &\quad \norm{q_{avoid}-q_p}^2 
\end{equation}
where $\cE^{c}$ is the set complement of the region $\cE$.

@@

  \subsubsection{Projection to the standard ellipse}\rtron{Move to the appendix or refer to the conference paper}
  The ellipse $\cE$ expressed in $\cF_\cE$ is given by $\Eframe{\cE}=\{\Eframe{q}\in\real{m}:d(q^\cE_1,\Eframe{q})+d(\Eframe{q},q^\cE_2)<2a\}$, where the coordinates of the two foci $q^\cE_1,q^\cE_2$ in $\cF_\cE$ are
  \begin{align}
    q^\cE_1=\bmat{c & 0 & 0}\transpose, && q^\cE_2=\bmat{-c & 0 & 0}\transpose,
  \end{align}
  with $c=\frac{\norm{q_2-q_1}}{2}$.

@@

For a intermediate step to get the relationship between a polygon shaped forbidden region, we first consider a forbidden region in the shape of a hyperplane $\mathcal{L}(\tilde{q}) = \{\tilde{q}\in\mathbb{R}^n: \vn\transpose \tilde{q} = d\}$ (as shown in Figure \ref{fig:Ellipse-to-plane}). The reachability constraint now can then be defined as $\mathcal{L} \cap \mathcal{E}(q_1,q_2,a) = \emptyset$.
  Using the transformation introduced in (\ref{sec:rotation2Standard}), the ellipse can be transferred into a standard one with the new hyperplane in the form of $\Eframe \cL(\tilde{q}) = \{  \tilde{q}\in\mathbb{R}^m:  \vn_\cE\transpose \tilde{q} = d_{\cE}\}$, where,
  \begin{equation}
    n_\cE = H(q_1,q_2)n
\end{equation}
\begin{equation} 
	d_\cE =-\vn\transpose o + d
  \end{equation}
\subsubsection{Projection via the tangent interpolation point}
  For each $\cL_\cE$, there exist two planes $\Eframe \cL_{1} = \{\tilde{q}\in\mathbb{R}^m: \vn_\cE \transpose \tilde{q} =  d_{\cE t} \}$ and $\Eframe \cL_{2} = \{\tilde{q}\in\mathbb{R}^m: \vn_\cE \transpose \tilde{q} = - d_{\cE t} \}$ that are parallel to $\cL$ and tangent to the ellipse, where
  \begin{equation}
    d_{\cE t} = \sqrt{ n_\cE\transpose Q^{-1} n_\cE}.
  \end{equation}
  The corresponding tangent points are given as 
  \begin{equation}
    p^\cE_{1} = \frac{  d_{\cE t} Q^{-1}   n_\cE}{ n_\cE \transpose Q^{-1}  n_\cE} = \frac{Q^{-1} n_\cE}{ d_{\cE t}},
  \end{equation}
  \begin{equation}
   p^\cE_{2} = -  p^\cE_{1}.
  \end{equation}
  \rtron{Refer to figure}
  We use these two tangent points to define a novel measure of displacement between a plane and an ellipse.
  \begin{definition}
    The \emph{tangent interpolation point} is defined as 
    \begin{equation}\label{equ:ProjectPoint}
      p^\cE_{\cL} = \frac{ d_\cE Q^{-1} n_\cE}{ n_\cE\transpose Q^{-1} n_\cE}
    \end{equation}
  \end{definition}
  Note that $p^\cE_{\cL} \in \mathcal{L}$ and when  $d_\cE=d_{\cE t}$ or $d_\cE=-d_{\cE t}$, $p^\cE_{\cL}=p^\cE_{1}$ or $p^\cE_{\cL}=p^\cE_{2}$, respectively. 
  % it is easy to see that $\Eframe p_{\cL} \in \mathcal{L}$, and $p_{\mathcal{L}}$ can be used to indicate the relationship between the $\mathcal{L}$ and $\mathcal{E}$, which is shown below. \zyang{how to emphaise this? maybe a graph here?}
  %The reason such that a specific point is used is for the convenience of evaluating ellipse-plane-segment relationship in Section~\ref{chapter:ellipse to segment}.
  It is clear that when $ d_\cE \in [- d_{\cE t},  d_{\cE t}]$, the plane $\cL$ and the ellipsoid $\cE$ have at least one intersection, thus violating our desired reachability constraint. Using this fact, we define the following projection operator:
  \begin{equation}\label{equ:plane2ellipse}
    \pi^\cE_{\vn \cE}(q) =\begin{cases}
      p^\cE_{t1}-p^\cE_{\cL} &  d_\cE \in [0, d_{\cE t}], \\
      p^\cE_{t2}-p^\cE_{\cL} &  d_\cE \in [- d_{\cE t},0), \\
      {0} & \textrm{otherwise}.
    \end{cases}
  \end{equation}
  \subsubsection{ADMM constraints}
  Transforming the projection $\pi^\cE$ back to $\cF$ we have:
  \begin{equation}\label{equ:constraint plane to ellipse}
    D_{\vn}(q)= H^{-1}(q(t_1),q(t_2))\pi^\cE_{\vn \cE}(q)+o,
  \end{equation}
  with the corresponding constraint written as
  \begin{equation}
    \zeta_\vn = \{q\in\mathbb{R}^{nm}:\norm{D_{\vn}(q)} = 0 \}
  \end{equation}
  \begin{equation}
   \quad\Pi_\vn(D_{\vn}(q)) = {0}
  \end{equation}

@@

\begin{figure}[htbp]
\begin{center}
\includegraphics[width=0.6\linewidth]{Ellipse2safezone_V2}
\caption{Polygon-ellipse constraint. This constraint is simplified to either a plane-ellipse constraint or a point-ellipse constraint.}
\label{fig:Ellipse-to-safezone}
\end{center}
\end{figure}
